The yaq Project: \\ Standardized Software Enabling Flexible Instrumentation

### 2023-02-03 paper.tex
cite
address many motors

The third hardware interface barrier: Some hardware interfaces are blocking.
Monolithic orchestration software often necessitates separate threads for each component hardware interface, a fragile pattern in which small mistakes become both critical and elusive errors.
In \yaq{}, each daemon is only responsible for a single hardware interface.
%Each \yaq{} daemon runs as a separate process (possibly not even on the same machine) and thus many of the hard concurrency problems with communicating with hardware are mitigated.
It is expected that each message call over the \yaq{} interface will return rapidly, ensuring that client applications are not blocked for extended periods of time.
This principle applies even when the message starts an action that might take several seconds to complete, such as homing a motor or initiating a measurement for a sensor.
In these instances, the initial message simply starts the action and returns, with a separate message provided to retrieve results when they exist.
In order to know how long to wait, the ``is-daemon`` trait provides a message called ``is\_busy'', which returns "true" while the long running action is not complete, and "false" once it is finished.
Additionally, multiple clients can communicate with the same daemon simultaneously.
A complex instrument may involve multiple operators watching sensor data in real time, while one program is orchestrating the hardware and recording the data.

@@

Open source software built by scientists often has incomplete and inadequate documentation.
\yaq{} attempts to automate daemon documentation as much as possible.
Our website, https://yaq.fyi, automatically builds generated reference pages for all known protocols.
These pages are automatically updated when new versions are published.
Our website also contains written and video tutorials on \yaq{} usage and development.

>>>>

The third hardware interface barrier: Some hardware interfaces are blocking or slow.
Experimental orchestration often means doing several things at once, \textit{e.g.} simultaneously moving several motors.
To accomplish this task performantly, monolithic orchestration software often necessitates separate threads for each component hardware interface, a fragile pattern in which small mistakes become both critical and elusive errors.
When writing orchestration software using \yaq{}, slow hardware interfaces are replaced by fast Avro-RPC.
In our experience, the \yaq{} interface is responsive enough to give good performace when orchestrating tens of motors using a singly-threaded client.
There is a longer discussion of timing and order-of-operations control details using \yaq{} in the Supplementary Material.

@@

Software built by scientists often has incomplete and inadequate documentation. \cite{SegalJudith2007a}
\yaq{} attempts to automate daemon documentation as much as possible.
Our website, https://yaq.fyi, automatically builds generated reference pages for all known protocols.
These pages are automatically updated when new versions are published.
Our website also contains written and video tutorials on \yaq{} usage and development.
